Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-27 ms.tex
cutting fat - and lots of it5

\begin{figure}
\includegraphics[width=\columnwidth]{vwSources.pdf}
\caption{\label{fig:vwSources} {\it Top Panel:} Sources contributing
  to the total heating rate of the CNM, $\vwO$ ({\it black}): stellar
  wind heating ({\it orange}), Ia supernovae ({\it green}),
  millisecond pulsars ({\it blue}), and compton heating ({\it pink}).
  Each heating source varies with black hole mass $\Mbh$ calculated
  for average SFHs from \citet{MosterNaab+:2013a}; see
  Appendix~\ref{app:windheat} for details.  {\it Bottom Panel:} The
  ratio of the total heating rate ($\dot{q}_{\rm heat} \propto
  v_{w}^{2}$) from the top panel to the radiative cooling rate
  ($|\dot{q}_{\rm rad}|$) at the stagnation radius (red), parameter
  $\eta$ characterizing stellar mass loss rate (eq.~[\ref{eq:q}]) as a
  function of black hole mass $\Mbh$, calculated for average star SFHs
  from \citealt{MosterNaab+:2013a} (blue), the ratio of Ia radius to
  stagnation radius as a function of $\Mbh$ (green), and the ratio of
  the normalized heating rate
  $\zeta=\sqrt{1+\frac{v_w^2}{\sigma_0^2}}$ to the critical heating
  rate $\zeta_c$ for which no outflow is possible (see
  eq.~\eqref{eq:zetacrit}). In evaluating the latter quantity we use
  scaling relations for $r_b$, $\rinf$, and $\Gamma$ based on
  \citet{LauerFaber+:2007a}. {\bf AG added the zeta/zetac line}.}

@@

\subsubsection{Continuous SFH}
A single burst of star formation does not describe the typical star
formation history of most galaxies.  Qualitatively, smaller galaxies
will on average have more recent star formation, resulting in
energetic young stellar winds and supernovae which dominate the gas
heating budget.  Massive galaxies, on the other hand, will on average
possess older stellar populations, with their heating rates dominated
by SN Ia (on large radial scales) and SMBH feedback.  We estimate the
average value of $\vwO$ as a function of SMBH mass for each heating
source by calculating its value using the average cosmic star
formation histories of \citet{MosterNaab+:2013a}.  Appendix
\ref{app:windheat} describes how the average SFH is
used to determine the stellar wind heating $v_{\rm w}^{\star}$ and
mass return parameter $\eta$ as a function of $M_{\bullet}$
(Fig.~\ref{fig:vwSources}; top panel).  SFHs are also
convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average SFH of galaxies containing a given black hole mass.  The
younger stellar populations characterizing galaxies with $\Mbh\lsim
3\times 10^{8} \Msun$ are dominated by stellar winds, with
$v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for $\Mbh\gsim 3\times
10^{8} \Msun$ does the lack of young stellar populations significantly
reduce the role of stellar wind heating.  But for these massive
galaxies, the Ia radius is sufficiently small for SN Ia to contribute
a comparable level of heating, $v_{w}^{\rm Ia} \gtrsim 600$ km
s$^{-1}$ (in low mass galaxies SN Ia do not contribute to the total
heating rate because $r_{\rm Ia} > \rs$; bottom panel).  Strikingly,
we find that galactic nuclei experiencing the same average SFH as
their host galaxies produce thermally stable flows across almost all
black hole masses.

An important caveat, however, is that $\vw$ is only a few times higher
than the stellar velocity dispersion, i.e. $\zeta \sim \zeta_c$
(eq.~[\ref{eq:zetacrit}]).  This implies that the true stagnation
radius and SMBH accretion rate could be larger than our analytic
estimates, potentially resulting in thermal instability, especially
for core galaxies with large break radii ($\S\ref{sec:mdot}$).  In
other words, the CNM of some subset of massive ellipticals be
thermally unstable, not necessarily because they are receiving lower
stellar feedback than the average for their galaxy mass, but due to
the high heating required for stability given the structure of their
stellar potential.

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{vwSources.pdf}
\caption{\label{fig:vwSources} {\it Top Panel:} Sources contributing
  to the total heating rate of the CNM, $\vwO$ ({\it black}): stellar
  wind heating ({\it orange}), Ia supernovae ({\it green}),
  millisecond pulsars ({\it blue}), and compton heating ({\it pink}).
  Each heating source varies with black hole mass $\Mbh$ calculated
  for average SFHs from \citet{MosterNaab+:2013a}; see
  Appendix~\ref{app:windheat} for details.  {\it Bottom Panel:} The
  ratio of the total heating rate ($\dot{q}_{\rm heat} \propto
  v_{w}^{2}$) from the top panel to the radiative cooling rate
  ($|\dot{q}_{\rm rad}|$) at the stagnation radius (red), parameter
  $\eta$ characterizing stellar mass loss rate (eq.~[\ref{eq:q}]) as a
  function of black hole mass $\Mbh$, calculated for average star SFHs
  from \citealt{MosterNaab+:2013a} (blue), the ratio of Ia radius to
  stagnation radius as a function of $\Mbh$ (green), and the ratio of $\zeta/\zeta_{c}$ (blue), where $\zeta_c$ (eq.~[\ref{eq:zetacrit}]) is the critical heating parameter $\zeta \equiv \sqrt{1+(v_w/\sigma_0)^2}$ below which outflows are impossible.  The latter quantity is evalulated using mean scaling relations for $r_b$, $\rinf$, and $\Gamma$ from the \citet{LauerFaber+:2007a} galaxy sample.}

@@

\subsubsection{Continuous SFH}
A single burst of star formation does not describe the typical star
formation history of most galaxies.  Qualitatively, smaller galaxies
will on average experience more recent star formation, resulting in
energetic young stellar winds and supernovae which dominate the gas
heating budget.  Massive galaxies, on the other hand, will on average
possess older stellar populations, with their heating rates dominated
by SN Ia (on large radial scales) and SMBH feedback.  We estimate the
average value of $\vwO$ as a function of SMBH mass for each heating
source by calculating its value using the average cosmic star
formation histories of \citet{MosterNaab+:2013a}.  Appendix
\ref{app:windheat} describes how the average SFH is
used to determine the stellar wind heating $v_{\rm w}^{\star}$ and
mass return parameter $\eta$ as a function of $M_{\bullet}$
(Fig.~\ref{fig:vwSources}; top panel).  SFHs are also
convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average SFH of galaxies containing a given black hole mass.  The
younger stellar populations characterizing low mass galaxies with $\Mbh\lsim
3\times 10^{8} \Msun$ are dominated by stellar winds, with
$v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for $\Mbh\gsim 3\times
10^{8} \Msun$ does the lack of young stellar populations significantly
reduce the role of stellar wind heating.  For these massive
galaxies, however, the Ia radius is sufficiently small that SN Ia contribute
a comparable level of heating, $v_{w}^{\rm Ia} \gtrsim 500$ km
s$^{-1}$ (in low mass galaxies SN Ia do not contribute to the total
heating rate because $r_{\rm Ia} > \rs$; bottom panel).

Strikingly, we find that galactic nuclei that experience the same average SFH as
their host galaxies possess thermally stable flows across all
black hole masses.  An important caveat, however, is that $\vw$ is generally only a few times higher than the stellar velocity dispersion, i.e. $\zeta \sim \zeta_c$
(eq.~[\ref{eq:zetacrit}]).  This implies that the true stagnation radius and SMBH accretion rate could be larger than our analytic estimates, potentially resulting in thermal instability in a significant fraction of galaxies.  The bottom panel of Figure~\ref{fig:vwSources} also shows $\zeta/\zeta_c$, where $\zeta_c$ is estimated from the best-fit power-law dependence of $r_b$, $\rinf$, and $\Gamma$ on $M_{\bullet}$ from the \citet{LauerFaber+:2007a} galaxy sample.   In other words, the CNM of a significant fraction of massive ellipticals may be thermally unstable, not necessarily because they are receiving lower stellar feedback than the average for their galaxy mass, but due to the high heating required for stability given the structure of their stellar potential.
